---
title: "CMNN Burden For Ghana, Kenya & Zimbabwe"
author: "Jerome Milimu"
output: github_document
---
# Communicable, Maternal, Neonatal, and Nutritional Disease Burden (CMNN)

```{r include=FALSE}
# Load necessary packages
pacman::p_load(tidyverse, 
               knitr, 
               here,
               janitor)

# Import the communicable diseases data
data_cmnn <- read_csv(here("data", "burden-of-disease-cmnn.csv"))
```

## Table of Estimates for CMNN Burden Over Time

```{r echo=FALSE}
# Here render a table for the DALY burden over time for the three countries 
# You should pivot the data to show each country in a separate column.
# Use kable() from the knitr package to render the table.

data_cmnn <- 
  data_cmnn %>% 
    clean_names()

#names(data_cmnn)

gha_ken_zwe_data_cmnn <-
  data_cmnn %>% 
    filter(entity  %in% c("Ghana",
                          "Kenya",
                          "Zimbabwe")) %>% 
    select(-code) %>% 
    rename(country = "entity",
           DALYs = "dal_ys_disability_adjusted_life_years_communicable_maternal_neonatal_and_nutritional_diseases_sex_both_age_age_standardized_rate")
  
#head(gha_ken_zwe_data_cmnn)


gha_ken_zwe_data_cmnn_wide <- 
  gha_ken_zwe_data_cmnn %>% 
    pivot_wider(names_from = "country",
                values_from = "DALYs") 

#head(gha_ken_zwe_data_cmnn_wide)

# Render table
 knitr::kable(gha_ken_zwe_data_cmnn_wide)


```



## Plot Showing Trends in CMNN Burden Over Time

```{r echo=FALSE}
gha_ken_zwe_data_cmnn %>% 
  ggplot(
    mapping = aes(x = year,
                  y = DALYs,
                  colour = country)) +
  theme_minimal() +
  geom_line() +
  labs(
    title = "Trends in CMNN Burden Over Time",
    subtitle = "Ghana, Kenya & Zimbabwe"
  )
```

## Summary of CMNN Burden Findings

Over the Period examined, Zimbabwe had the highest DALYs related to communicable, maternal, neonatal, and nutritional disease burden. Followed by Kenya and finally Ghana. It was interesting to note that the DALYs followed a similar trend to the burden of the HIV/AIDS pandemic faced by each country 


## Cumulative DALY Burden For The Three Countries Over Time

```{r echo=FALSE}
# Calculate cumulative DALYs
gha_ken_zwe_data_cmnn <- 
  gha_ken_zwe_data_cmnn %>%
  group_by(country) %>%
  arrange(year) %>%
  mutate(Cumulative_DALYs = cumsum(DALYs))

# Plotting the cumulative DALY burden
ggplot(gha_ken_zwe_data_cmnn, 
       aes(x = year, 
           y = Cumulative_DALYs, 
           color = country)) +
  geom_line() +
  labs(title = "Cumulative DALY Burden Over Time by Country",
       x = "Year",
       y = "Cumulative DALYs",
       color = "Country") +
  theme_minimal()
```

