---
title: "Overall Disease Burden for Ghana, Kenya & Zimbabwe"
author: "John Andoh"
output: github_document
---

# Overall Disease Burden

We investigate the overall disease burden in DALYs among three countries; Ghana, Kenya, and Zimababwe

```{r echo=F, message=F}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the overall disease data
data_overall <- read_csv(here("data", "burden-of-disease-all-causes.csv"))
```

## Table of Estimates for Overall Disease Burden Over Time

```{r echo=F, message=F}
disease_burden_table <- data_overall %>% 
  filter(Entity %in% c("Ghana", "Kenya", "Zimbabwe")) %>% 
  select(-Code) %>% 
  rename(country = "Entity",
         DALYs = `DALYs (Disability-Adjusted Life Years) - All causes - Sex: Both - Age: Age-standardized (Rate)`) %>% 
  pivot_wider(names_from = "country", 
              values_from = DALYs) %>% 
  kable()
disease_burden_table  

# Here render a table for the DALY burden over time for the three countries 
# You should pivot the data to show each country in a separate column.
# Use kable() from the knitr package to render the table.
```

## Plot Showing Trends in Overall Disease Burden Over Time

```{r, echo=F, message=F}

filtered_data <- data_overall %>%
  filter(Entity %in% c("Kenya", "Zimbabwe", "Ghana"))

# Plotting the line graph with ggplot2
disease_burden_plot <- ggplot(filtered_data, aes(x = Year, y = `DALYs (Disability-Adjusted Life Years) - All causes - Sex: Both - Age: Age-standardized (Rate)`, group = Entity, color = Entity)) +
  geom_line() +
  labs(title = "Trend of DALYs Burden Over Time",
       x = "Year",
       y = "DALYs (Rate)",
       color = "Country") +
  theme_minimal()
disease_burden_plot
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.
```

## Summary of Overall Disease Burden Findings

Provide a brief analysis based on the data presented in the table and chart. Highlight any significant findings or patterns. About 3 sentences.

Zimbabwe had the highest overall disease burden among the three countries over time. All countries experienced a peak in disease burden the early 2000s and had been in a sharp decline since then. This can be attributed to the massive global health efforts in recent years.
